---
title: "Predicting Bitcoin price and volume by hashrate changes" 
author: "Brian Kim"
output: html_document
---

```{=html}
<style>
body {
text-align: justify}
</style>
```
```{r setup, include = FALSE, out.width = "100%", out.height = "100%"}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ezids)
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
library(ggfortify)  # for plot_acf
library(tidyverse)
library(lubridate)
library(fpp2)
library(astsa)
library(cowplot)
#options(repos = c(CRAN = "https://cloud.r-project.org/"))
```

```{r}
# Load data
df <- read_csv("/Users/brian/Documents/GitHub/Bitcoin_mining_PJT2_Only_Brian/Cleaned_bitcoin_mining_price.csv")
summary(df)
```

```{r}
# Display the structure of your data
str(df)
```

# Data Preprocessing
```{r}
# missing value check
missing_values <- sum(is.na(df))
missing_values_per_variable <- colSums(is.na(df))
print(missing_values)
print(missing_values_per_variable)

# Replace missing values
library(dplyr)

# Replace missing values with the median for each variable in the dataframe df
df_filled <- df %>%
  mutate_all(funs(ifelse(is.na(.), median(., na.rm = TRUE), .)))


```

# Time series analys of bitcoin price

```{r}

#conver data format
df$Date <- as.Date(df$Date, format = "%m/%d/%Y")

# bitcoin price
ggplotly(ggplot(df, aes(Date, BTC_Price)) + geom_line(col = 'gold1') + 
  labs(title = 'Bitcoin Price', x = 'Year') +
  scale_y_continuous(breaks = c(0, 5000, 10000, 15000, 25000, 35000, 45000, 55000, 65000), 
                     labels = c('$0', '$5,000', '$10,000', '$15,000', '$25,000','$35,000','$45,000','$55,000','$65,000')))



```

```{r}
# Convert date format
df$Date <- as.Date(df$Date, format = "%m/%d/%Y")

# Create Bitcoin price plot
bitcoin_plot <- ggplot(df, aes(Date, BTC_Price)) + 
  geom_line(col = 'gold1') + 
  labs(title = 'Bitcoin Price', x = 'Year') +
  scale_y_continuous(
    name = "BTC_Price",
    breaks = c(0, 5000, 10000, 15000, 25000, 35000, 45000, 55000, 65000), 
    labels = c('$0', '$5,000', '$10,000', '$15,000', '$25,000','$35,000','$45,000','$55,000','$65,000')
  )

# Create Hashrate plot
hashrate_plot <- ggplot(df, aes(Date, Hashrate)) + 
  geom_line(col = 'blue') + 
  labs(title = 'Hashrate', x = 'Year') +
  scale_y_continuous(
    name = "Hashrate",
    sec.axis = sec_axis(trans = ~ . * 1, 
                        breaks = c(50000, 100000, 150000),
                        labels = c('50k', '100k', '150k'))
  )

combined_plot <- plot_grid(bitcoin_plot, hashrate_plot, align = "v", ncol = 1)

# Display the combined plot
print(combined_plot)

```

# This analyis should be deleted

```{r}

# This analyis should be deleted

# convert date format
df$Date <- as.Date(df$Date, format = "%m/%d/%Y")

bit_ts = df %>%
  filter(Date > as.Date('1/1/2017')) %>%
  arrange(Date) %>%
  select(BTC_Price) %>%
  as.matrix() %>%
  ts()

gglagplot(bit_ts, do.lines = F) +
  scale_color_continuous(low = "#b37400", high = "#ffc04d", 
                         breaks = c(1, 1600, 2500, 3500, 4500, 27500, 65000), 
                         labels = c('2017', '2018', '2019', '2020', '2021', '2022', '2023')) + 
  scale_y_continuous(breaks = c(0, 15000, 40000, 65000), 
                     labels = c('$0', '$15,000', '$40,000', '$65,000')) +
  scale_x_continuous(breaks = c(5000, 40000, 65000), 
                     labels = c('$5,000', '$40,000', '$65,000')) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10),
    axis.text.x = element_text(size = 6),
    axis.text.y = element_text(size = 6),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10)
  )

```

```{r}
cut_bit_df = df %>%
  filter(Date > as.Date('2017-01-01'))

ggplotly(cut_bit_df[-1,] %>%
  mutate(BTC_Price = diff(cut_bit_df$BTC_Price)) %>%
  ggplot(aes(Date, BTC_Price)) + geom_line(col = '#ffa500') + 
  labs(x = '', title = 'Bitcoin Differenced By One', y = 'Difference'))


```

# Corrleation Analysis

```{r, fig.height = 12, fig.width = 12, fig.align = 'center'}
library(corrplot)

# Check the data types of each column
column_types <- sapply(df, class)

# Identify non-numeric columns
non_numeric_columns <- names(column_types[column_types != "numeric"])

# Convert non-numeric columns to numeric
df[, non_numeric_columns] <- lapply(df[, non_numeric_columns], as.numeric)

# corrleation matrix
correlation_matrix <- cor(df, use = "complete.obs")

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF",
                          "#77AADD", "#4477AA"))

corrplot(correlation_matrix,  # Replace with your correlation matrix
          method = "color",  # Use color to represent correlations
          col = col(200),  # Color palette
          type = "full",  # Display the upper triangle of the matrix
          order = "hclust",  # Order variables using hierarchical clustering
          addCoef.col = "black",  # Color of correlation coefficient labels
          tl.col = "black",  # Text label color
          tl.srt = 90,  # Text label rotation angle
          tl.cex = 0.8,
          sig.level = 0.01,  # Significance level
          insig = "blank",  # Display "blank" for insignificant correlations
          # Hide correlation coefficient on the principal diagonal
          diag = TRUE
)
```

# Linear regression model for bitcoin price by hashrate changes

```{r}
# Load required libraries
library(tidyverse)
library(caret)

# Convert date format
df_filled$Date <- as.Date(df$Date, format = "%m/%d/%Y")

# Feature engineering
df_filled$Year <- as.numeric(format(df_filled$Date, "%Y"))
df_filled$Month <- as.numeric(format(df_filled$Date, "%m"))
df_filled$Day <- as.numeric(format(df_filled$Date, "%d"))

# Split data into training and testing sets
# set.seed(123)
# train_index <- createDataPartition(df$Price, p = 0.8, list = FALSE)
# train_data <- df[train_index, ]
# test_data <- df[-train_index, ]

# Define predictor variables
predictors <- c("Year", "Month", "Day", "Hashrate")

# Train a predictive model (linear regression for simplicity)
model <- lm(BTC_Price ~ ., data = train_data[, c(predictors, "BTC_Price")])

# Make predictions on the test set
predictions <- predict(model, newdata = test_data)

# Visualize predictions vs. actual prices
plot(test_data$Date, test_data$BTC_Price, col = "blue", type = "l", lty = 1, xlab = "Date", ylab = "BTC_Price")
lines(test_data$Date, predictions, col = "red", type = "l", lty = 2)
legend("topright", legend = c("Actual", "Predicted"), col = c("blue", "red"), lty = 1:2)


# Evaluate the model
rmse <- sqrt(mean((predictions - test_data$Price)^2))
cat("Root Mean Squared Error(RMSE) for BTC_Price:", rmse, "\n")

```

# Linear regression model for bitcoin volume by hashrate changes

```{r}
# Load required libraries
library(tidyverse)
library(caret)

# Convert date format
df_filled$Date <- as.Date(df$Date, format = "%m/%d/%Y")

# Feature engineering
df_filled$Year <- as.numeric(format(df_filled$Date, "%Y"))
df_filled$Month <- as.numeric(format(df_filled$Date, "%m"))
df_filled$Day <- as.numeric(format(df_filled$Date, "%d"))

# Split data into training and testing sets
# set.seed(123)
# train_index <- createDataPartition(df$Volume, p = 0.8, list = FALSE)
train_data <- df_filled[train_index, ]
test_data <- df_filled[-train_index, ]
Testnewdata <- test_data[, !(names(test_data) %in% c("BTC_Price"))]

# Define predictor variables
predictors <- c("Year", "Month", "Day", "Hashrate")

# Train a predictive model (linear regression for simplicity)
model <- lm(BTC_Volume ~ ., data = train_data[, c(predictors, "BTC_Volume")])

# Make predictions on the test set
predictions <- predict(model, newdata = test_data)

# Visualize predictions vs. actual prices
plot(test_data$Date, test_data$Volume, col = "blue", type = "l", lty = 1, xlab = "Date", ylab = "Volume")
lines(test_data$Date, predictions, col = "red", type = "l", lty = 2)
legend("topright", legend = c("Actual", "Predicted"), col = c("blue", "red"), lty = 1:2)

# Evaluate the model
rmse <- sqrt(mean((predictions - test_data$Volume)^2))
cat("Root Mean Squared Error(RMSE) for BTC_Volume:", rmse, "\n")
```

# RandomForest model for predicting bitcoin price by hashrate changes

```{r}
# Load required libraries
library(tidyverse)
library(caret)
library(randomForest)

# Convert date format
df_filled$Date <- as.Date(df_filled$Date, format = "%m/%d/%Y")

# Feature engineering
df_filled$Year <- as.numeric(format(df_filled$Date, "%Y"))
df_filled$Month <- as.numeric(format(df_filled$Date, "%m"))
df_filled$Day <- as.numeric(format(df_filled$Date, "%d"))

# Split data into training and testing sets
# set.seed(123)
# train_index <- createDataPartition(df$Price, p = 0.8, list = FALSE). #time series split data
train_data <- df_filled[train_index, ]
test_data <- df_filled[-train_index, ]
Testnewdata <- test_data[, !(names(test_data) %in% c("BTC_Price"))]

# Define predictor variables
predictors <- c("Silver_Price", "SP500_Price", "ETH_Price", "Hashrate")

# Train a model for predicting volume (using Random Forest for simplicity)
Price_model <- randomForest(BTC_Price ~ ., data = train_data[, c(predictors, "BTC_Price")])

# Make predictions on the test set for volume
price_predictions <- predict(Price_model, newdata = Testnewdata)

# Visualize predictions vs. actual volume
plot(Testnewdata$Date, Testnewdata$BTC_Price, col = "blue", type = "l", lty = 1, xlab = "Date", ylab = "BTC_Price")
lines(Testnewdata$Date, price_predictions, col = "red", type = "l", lty = 2)
legend("topright", legend = c("Actual", "Predicted"), col = c("blue", "red"), lty = 1:2)

# Combine dates and predicted prices into a data frame
predicted_data <- data.frame(
  Date = Testnewdata$Date,
  Predicted_Price = price_predictions
)

# Print the predicted data
print(predicted_data)

# Evaluate the volume model
rmse_price <- sqrt(mean((price_predictions - Testnewdata$BTC_Price)^2))
cat("Root Mean Squared Error(RMSE) for BTC_Price:", rmse_price, "\n")

print(train_data[, c(predictors, "BTC_Price")])
print(colnames(Testnewdata))
```

# RandomForest model for predicting bitcoin volume by hashrate changes

```{r}
# Load required libraries
library(tidyverse)
library(caret)
library(randomForest)

# Convert date format
df_filled$Date <- as.Date(df_filled$Date, format = "%m/%d/%Y")

# Feature engineering
df_filled$Year <- as.numeric(format(df_filled$Date, "%Y"))
df_filled$Month <- as.numeric(format(df_filled$Date, "%m"))
df_filled$Day <- as.numeric(format(df_filled$Date, "%d"))

# Split data into training and testing sets
# set.seed(123)
# train_index <- createDataPartition(df$Price, p = 0.8, list = FALSE)
train_data <- df_filled[train_index, ]
test_data <- df_filled[-train_index, ]
Testnewdata <- test_data[, !(names(test_data) %in% c("BTC_Volume"))]

# Define predictor variables
predictors <- c("Silver_Price", "SP500_Price", "ETH_Price", "Hashrate")

# Train a model for predicting volume (using Random Forest for simplicity)
volume_model <- randomForest(BTC_Volume ~ ., data = train_data[, c(predictors, "BTC_Volume")])

# Make predictions on the test set for volume
volume_predictions <- predict(volume_model, newdata = Testnewdata)

# Visualize predictions vs. actual volume
plot(Testnewdata$Date, Testnewdata$BTC_Volume, col = "blue", type = "l", lty = 1, xlab = "Year", ylab = "BTC_Volume")
lines(Testnewdata$Date, volume_predictions, col = "red", type = "l", lty = 2)
legend("topright", legend = c("Actual", "Predicted"), col = c("blue", "red"), lty = 1:2)

# Combine dates and predicted prices into a data frame
predicted_data <- data.frame(
  Date = test_data$Date,
  Predicted_Volume = volume_predictions
)

# Print the predicted data
print(predicted_data)

# Evaluate the volume model
rmse_volume <- sqrt(mean((volume_predictions - Testnewdata$Volume)^2))
cat("Root Mean Squared Error(RMSE) for Volume:", rmse_volume, "\n")

print(train_data[, c(predictors, "BTC_Volume")])
print(colnames(Testnewdata))

```
