---
title: "DS_EDA PJT1"
author: "Brian Kim"
date: "`r Sys.Date()`"
output:
  html_document:
  code_folding: hide
number_sections: false
toc: yes
toc_depth: 3
toc_float: yes
pdf_document:
  toc: yes
---
  
```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session
```

library(ezids)
library(readr) #install & load 
library(ggplot2)
library(DBI) #SQL
library(magrittr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(graphics) 
library(gridExtra)


```{r}
df <- read.csv("Regional harshrate global 1.csv")
dim(df)
str(df)
```

# monthly_hashrate boxplot
```{r}
boxplot(df$monthly_absolute_hashrate_EH.S)
```

# data grouping by country
```{r}
library(dplyr)

df %>%
  group_by(country) %>%
  summarise(count = n(), count_unique = n_distinct(country))
```

# Data extract
```{r}
# library(dplyr)
# df_us <- df %>%
#  filter(country == 'United States') %>%
#df_rs <- df %>%
#  filter(country == 'Russian Federation') %>%
# df_ca <- df %>%
#  filter(country == 'Canada') %>%
# df_ga <- df %>%
#  filter(country == 'Germany *') %>%
# df_ir <- df %>%
#  filter(country == 'Iran, Islamic Rep.') %>%
# df_il <- df %>%
#  filter(country == 'Ireland *') %>%
# df_ka <- df %>%
#  filter(country == 'Kazakhstan') %>%
# df_ch <- df %>%
#  filter(country == 'Mainland China') %>%
# df_ma <- df %>%
#  filter(country == 'Malaysia') %>%
# df_ot <- df %>%
#   filter(country == 'Other') %>%
```


```{r}
# histogram
library(ggplot2)

ggplot(data = df, aes(x = monthly_absolute_hashrate_EH.S)) + 
  geom_histogram()

# histogram of each country
min(df$monthly_absolute_hashrate_EH.S)
max(df$monthly_absolute_hashrate_EH.S)
```

```{r}
# data merge
df1 <- read.csv("bitcoin_mining.csv")
str(df1)
```

```{r}
names(df1)[names(df1)=="Date.and.Time"] <-"date"
names(df1)[names(df1)=="Estimated..MtCO2e"] <-"MtCO2e"
names(df1)[names(df1)=="annualised.consumption.GUESS..TWh"] <-"annual_consumption"
df1$date <- gsub('T00:00:00', '', df1$date)

df <- merge(df,df1,by="date")
```

# Scatterplot of hashrate of regional Data
```{r}
df <- read.csv("Regional harshrate global 1.csv")
df$date <- as.Date(df$date)

df <- df %>%
  mutate(year = lubridate::year(date),  
         month = lubridate::month(date))

countries <- unique(df$country)
plot_list <- lapply(countries, function(country) {
  country_df <- df %>% filter(country == !!country)
  p <- ggplot(country_df, aes(x = as.Date(paste(year, month, "01", sep = "-")), y = monthly_absolute_hashrate_EH.S)) +
    geom_line(color = "blue") +
    geom_point(shape = 21, color = "black", fill = "red", size = 1) +
    labs(x = "Date", y = "Hashrate") +
    theme_minimal() +
    ggtitle(paste("Bit_Hashrate", country))
    
  # Create a plot title with a smaller size
  p <- p + theme(plot.title = element_text(size = 10))

  # Print the plot
  print(p)

  return(p)
})



```

```{r}
df <- read.csv("Regional harshrate global 1.csv")
df$date <- as.Date(df$date)

df <- df %>%
  mutate(year = lubridate::year(date),  
         month = lubridate::month(date))

countries <- unique(df$country)
plot_list <- lapply(countries, function(country) {
  country_df <- df %>% filter(country == !!country)
  p <- ggplot(country_df, aes(x = as.Date(paste(year, month, "01", sep = "-")), y = monthly_absolute_hashrate_EH.S)) +
    geom_line(color = "blue") +
    geom_point(shape = 21, color = "black", fill = "red", size = 1) +
    labs(x = "Date", y = "Hashrate") +
    theme_minimal() +
    ggtitle(paste("Bit_Hashrate", country)) +
    theme(plot.title = element_text(size = 10))

  return(p)
})

# Display the plots in a grid
library(gridExtra)
grid.arrange(grobs = plot_list, ncol = 2)

```




```{r}
library(ggplot2)
library(dplyr)
library(hrbrthemes)
library(viridis)

df <- df %>%
  mutate(year = lubridate::year(date),  
         month = lubridate::month(date))

# Summarize the data by country and year
summary_df <- df %>%
  group_by(country, year) %>%
  summarise(n = sum(monthly_absolute_hashrate_EH.S))

countries <- unique(summary_df$country)

summary_df %>%
  ggplot(aes(x = year, y = n, group = country, color = country)) +
  geom_line() +
  scale_color_viridis(discrete = TRUE) +
  ggtitle("Bit_Hashrate") +
  theme_ipsum() +
  ylab("Hashrate")
```



# geom map plot - practice
```{r}
library(tidyverse)
library(ggthemes)

world_map = map_data("world") %>% 
  filter(! long > 180)

countries = world_map %>% 
  distinct(region) %>% 
  rowid_to_column()

countries %>% 
  ggplot(aes(fill = rowid, map_id = region)) +
  geom_map(map = world_map) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  coord_map("moll") +
  theme_map()
```
# map visualization - practice
```{r}
library(ggiraphExtra)
library(ggplot2)

df2 <- read.csv("Regional harshrate global 1-2.csv")
class(df2$monthly_hashrate)
df2$monthly_hashrate <- as.numeric(df2$monthly_hashrate)
ggChoropleth(data = df2, aes(fill = monthly_hashrate, map_id = country), map = world_map)

```


# map visualization

```{r}
library(ggplot2)
library(ggmap)
library(raster)
library(rgeos)
library(maps)
library(knitr)
library(rgdal)
library(sf)
library(tidyverse)
library(ggthemes)

# Load map data
world_map <- map_data("world") %>%
  filter(!long > 180)

# Create a data frame to hold your CSV data
df3 <- read.csv("Regional harshrate global 1-3.csv")

# Rename the "country" column to "region" to match your map data
colnames(df3)[colnames(df3) == "country"] <- "region"

# Group and summarize the data
df3_pre <- df3 %>%
  group_by(region) %>%
  summarise(
    avg_monthly_hashrate_share = mean(monthly_hashrate_share),
    avg_monthly_hashrate = mean(monthly_hashrate)
  )

# Merge the map and CSV data
merge_result <- merge(world_map, df3_pre, by = "region", all.x = TRUE)

# Create a map visualization
ggplot(data = merge_result) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = avg_monthly_hashrate)) +
  expand_limits(x = merge_result$long, y = merge_result$lat) +
  coord_map("moll") +
  labs(fill = "avg_monthly_hashrate") +
  theme_map()



```

#Summary statistic of regional bitcoin mining data analsysis

```{r}
library(dplyr)

df2 <- read.csv("Regional harshrate global 1-2.csv")

# Define the classification based on the 'country' column
df2 <- df2 %>%
  mutate(classification = ifelse(country %in% c("United States", "Mainland China"), country, "Other"))

head(df2,n=5)


```

# Hypothesis test - Anova test

Null hypothesis: H0:
In the United States, China, and other countries, as Bitcoin transaction volume increases, hashrate and carbon dioxide emissions will increase equally.

Alternative hypothesis: H1:
The United States, China, and other countries will not all have the same hashrate and average carbon dioxide emissions.
The p-value is 0.0008, which is < α.
 *(0.05).
So we can reject H0
   The alternative hypothesis is supported at the 95% significance level.


```{r}
library(ezids)
na_df2 <- is.na(df2)
sum(na_df2)
```
```{r}
anova_result <- aov(monthly_hashrate ~ classification, data = df2)
anova_summary_df2 = summary(anova_result)
xkabledply(anova_summary_df2)
```

